---
title: "Analysing phonological systems:"
subtitle: "on Bayesian typological research"
author: "George Moroz"
institute: "NRU HSE, Linguistic convergency laboratory"
date: "24 August 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
theme_set(theme_minimal())
library(waffle)

set.seed(42)
all_langs <- tibble(group = "total = 7000,\nvalue 1 = 1794,\nθ = 0.256",
                    value = sample(c("value 1", "value 2"),
                                   7000,
                                   replace = TRUE,
                                   prob = c(0.25, 0.75)))
```

# In this talk I will try to cover:

* goals of linguistic typology
* role of the sampling
* strategy for avoiding sampling

---

# Goals of the linguistic typology?

--
* attest a typological distributions

--

* find correlation of different typological distributions
    * absolute universals
    * distributional patterns and tendencies
    * semantic maps
    * diachronic change of typological values
    
--

* correlation of linguistic and non-linguistic patterns
    * population movements
    * population size
    * language contact
    * sociolinguistic paramtres
    * geopolitical environment
    
--
  
* deal with a mixed typological value per language

---

In this talk I will cover only simple case of an abstract binary typological category with two values: *val_1* and *val_2*.

The main question: if I find a new language somwhere in the Caucasus, what is the probability $\theta$ that it will have value *val_1*?

--

The most common way of finding answer will be:

* make a sample of languages
* estimate probability $\hat{\theta}$ using obtained sample
--
* ... pretend that $\hat{\theta}$ is kind of estimation of $\theta$, for more convincingness add (binomial) confidence interval

---
# On sampling: stats

.pull-left[
1. randomly pick some languages
]

.pull-right[
```{r}
all_langs %>%
  sample_n(200) %>%
  mutate(group = "total = 200,\nvalue 1 = 49,\nθ = 0.245") %>%
  rbind(all_langs) %>%
  count(group, value) ->
  simple_sample_results

simple_sample_results %>%
  ggplot(aes(fill = value, values = n)) +
  geom_waffle(color = "white", size = .3, n_rows = 70, flip = TRUE, show.legend = FALSE) +
  facet_wrap(~group, nrow = 1, strip.position = "bottom")+
  ggthemes::scale_fill_tableau(name=NULL)+
  theme_minimal(base_family = "Roboto Condensed")+
  scale_x_discrete()+
  scale_y_continuous(labels = function(x) x * 70, # make this multiplyer the same as n_rows
                     expand = c(0,0))+
  theme(panel.grid = element_blank(), axis.ticks.y = element_line())
```

]


---
# On sampling: linguistics

---
# Ok, but we are on the phonological section!

* It is possible to use phonological units or relation of any phonological theory you like:
    * features, feet, syllables, etc.
    * feature constituents, OT constraints, examplars, phonological alternations
    * phonological destinctions (e. g. /i/ vs. /ɨ/)
